Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-05 WSSSPE1_report/WSSSPE1_summary_report.tex
Breaks lines after 70 chars in Policy.
This makes it much easier to read online (such as on Github), compared to if
paragraphs are just a single line.
Changes presentation of Maturity Models.
These changes are in response to Von Welch's comment in the Google Doc tha a
"maturity model for scientific software would be generally useful, with
security as one example."

%\todo{seems like definitions will be mentioned elsewhere in the paper?}
Each of the framework papers included a definition of sustainability, with many overlaps.  In essence the papers related the question to change over time, an idea well captured in the definition used by the UK's Scientific Software Sustainability Institute: ``software you use today will be available---and continue to be improved and supported---in the future,'' as quoted in~\cite{Venters_WSSSPE}. This idea was also expressed as software that continues to serve its users~\cite{Pierce_WSSSPE}.
See \S\ref{sec:cross-cutting} for more on the discussion of definitions of sustainability in the workshop.

@@

Another aspect of sustainable software that is rarely given due
diligence in scientific software design is software security. Heiland
et al~\cite{Heiland_WSSSPE} discuss how the security measures
appropriate for scientific software strongly depend on the maturity
level of the software. However, the maturity level of a software
project meant to be sustained changes tremendously over its
development life cycle, and the eventual maturity level is often
difficult to predict during initial development. Best practice models
for identifying and defining cybersecurity vulnerabilities at
different stages of the software life cycle have been formalized as
Software Security Maturity Models. Although these are widely used in
industry, awareness among scientific software development communities
is low. Developing such models for scientific software would align
well with the objective of providing implementation approaches that
the scientific community can actually follow: A Software Security
Maturity Model provides a path to tightening security practices as
software matures, emphasizes understandability over complexity, and
classifies cybersecurity-pertinent practices for easy recognition of
what level of cybersecurity a software addresses.

>>>>

% \todo{seems like definitions will be mentioned elsewhere in the
% paper?}
Each of the framework papers included a definition of sustainability,
with many overlaps.  In essence the papers related the question to
change over time, an idea well captured in the definition used by the
UK's Scientific Software Sustainability Institute: ``software you use
today will be available---and continue to be improved and
supported---in the future,'' as quoted in~\cite{Venters_WSSSPE}. This
idea was also expressed as software that continues to serve its
users~\cite{Pierce_WSSSPE}.  See \S\ref{sec:cross-cutting} for more on
the discussion of definitions of sustainability in the workshop.

@@

Many facets of engineering for software sustainability strongly depend
on the maturity level of the software. However, the maturity level of
a software project meant to be sustained changes tremendously over its
development life cycle, and the eventual maturity level is often
difficult to predict during initial development. Using software
security as their case sudy, Heiland et al~\cite{Heiland_WSSSPE}
discuss how Maturity Models can be used to formalize best practices
appropriate for the different levels of maturity that a software
project may go through over its lifecycle. Cybersecurity is also an
example of a sustainability-relevant aspect in software design that is
rarely given due diligence in science. In particular in industry,
cybersecurity best practices for different stages of life cycle and
maturity have been formalized as Software Security Maturity Models,
and are widely used, yet awareness of these among scientific software
development communities remains low. In providing a path to tightening
security practices as software matures, such models align with the
objective of providing implementation approaches that the scientific
community can actually follow.
